---
title: "Introduction to Programming"
author: 
    name: "Monique Hernandez"
    email: mahernandez@ucsb.edu
    affiliation: 
    - University of California, Santa Barbara
    - Masters of Environmental Data Science, 2027
    - B.S. in Environmental Studies, 2025
date: "Augustt 10, 2026"
toc: true
format:
  html:
    theme: "pulse"
    other-links:
      - text: Course Website EDS 221 
        href: https://eds-221-programming-m26.github.io/course-materials/day1.html
      - text: Monique's website
        href: moniqueallyson.github.io
execute:
  message: false
  warning: false
---

[click here for github link](https://github.com/moniqueallyson/eds221day1) 

## Comment

```{r}
 #This is a comment, it starts with a pound sign
```


## Expressions

```{r}
2024
coral <- "Porites"
print(coral)
sample(1:6, size = 1)
```

## Definitions
:::{.callout-note}
An **expression** is a complete, self-contained piece of code that R can evaluate.
We write code in a **script** (this document).
We run code at the **console** (panel below).
Code modifies the **environment** panel to the right.
:::

## Assign variables
```{r}
1:6
d6 <- 1:6
d6

```

::: {.callout-tip}
For code to modify the environment, we have to assign the output to a variable using <-
:::

## PEMDAS
```{r}

# What do you think these operators do?
5 + 2
3 * 2
5 > 2

# Math operators follow PEMDAS rules
1 + 2 * 3
(1+ 2) * 3
1 + (2 * 3)

# Run this code and decide if * or : has higher precedence
2 * 3 : 8
```

## Functions
::: {.callout-tip}
are _triggered by parentheses_
:::
```{r}
1:6
d6 <- 1:6
d6
sample(d6, size = 1)

# What do you think this does?
sum(sample(d6, size = 2, replace = TRUE))

# If you want to work with text instead of numbers, use quotations
coral_genus <- "Porites"
# To create multiple text values, use c()
coral_genera <- c("Porites", "Acropora")
coral_genera

```